---
title: ""
format: html
page-layout: full
editor: visual
execute:
  message: false
  warning: false
  echo: false
  cache: false
---

# Workshops

```{r}
# Load necessary libraries
library(readr)
library(dplyr)
library(tidyr)
library(reactable)
library(ggplot2)
library(maps)
library(htmltools)
library(tidygeocoder)
library(plotly)
library(countrycode)
library(lubridate)
library(stringr)
library(purrr)
```

Workshops using the [Bioconductor Carpentry training material](https://www.bioconductor.org/help/education-training/) are listed below. If you have taught a workshop using the Bioconductor Carpentry material, please submit a pull request to update the [workshops TSV here](https://github.com/Bioconductor/bioconductor-training/blob/devel/data/workshops.tsv).

If you would like to request a Bioconductor Carpentry workshop at your institution, please fill out our [workshop request form](https://forms.gle/PGGRqFbmLrzTQmTY7).

See [this page](instructors.qmd) for Carpentry instructors in the Bioconductor community.

```{r}
# Function to determine workshop status and return parsed date
determine_status_and_date <- function(date_str) {
  parsed_date <- NA
  base_status <- "upcoming"

  if (grepl("TBD|Early|Late", date_str, ignore.case = TRUE)) {
    base_status <- "upcoming"
    # Use a future date for sorting, but one that reflects the year if present
    year_match <- str_extract(date_str, "\\d{4}")
    if (!is.na(year_match)) {
        parsed_date <- as.Date(paste0(year_match, "-12-31"))
    } else {
        parsed_date <- as.Date("2099-12-31") # Far future for TBD without year
    }
  } else {
    # Attempt to parse the date
    if (grepl("-\\s*\\d{1,2}\\s+[A-Za-z]+\\s+\\d{4}", date_str)) {
      end_date_str <- str_extract(date_str, "\\d{1,2}\\s+[A-Za-z]+\\s+\\d{4}$")
      parsed_date <- dmy(end_date_str, quiet = TRUE)
    } else if (grepl("\\d{1,2},\\s*\\d{1,2},\\s*\\d{1,2},\\s*\\d{1,2}\\s+[A-Za-z]+\\s+\\d{4}", date_str)) {
      last_date_part <- str_extract(date_str, "\\d{1,2}\\s+[A-Za-z]+\\s+\\d{4}$")
      parsed_date <- dmy(last_date_part, quiet = TRUE)
    } else if (grepl("^[0-9]{4}$", date_str)) {
      parsed_date <- ymd(paste0(date_str, "-12-31"), quiet = TRUE)
    } else {
      parsed_date <- dmy(date_str, quiet = TRUE)
      if (is.na(parsed_date)) {
        parsed_date <- my(date_str, quiet = TRUE)
      }
    }

    if (is.na(parsed_date)) {
      base_status <- "upcoming"
      parsed_date <- as.Date("2099-12-31")
    } else {
      base_status <- if (parsed_date >= today()) "upcoming" else "past"
    }
  }
  
  return(list(status = base_status, parsed_date = as.Date(parsed_date)))
}

# Add 'region' using countrycode for World Bank regions
workshops_raw <- read_tsv("../../data/workshops.tsv") %>%
  filter(initiative == "Carpentries")

# Apply the function and extract results
status_and_date_results <- mapply(determine_status_and_date, workshops_raw$date, SIMPLIFY = FALSE)

workshops <- workshops_raw %>%
  mutate(
    status = map_chr(status_and_date_results, "status"),
    sort_date = map(status_and_date_results, "parsed_date") %>% unlist() %>% as.Date()
  ) %>%
  mutate(Location = paste(city, country, sep = ", ")) %>%
  mutate(region = countrycode(country, origin = "country.name", destination = "region")) %>%
  arrange(desc(status), desc(sort_date))

# Filter out workshops that are "Online" for mapping purposes only
workshops_for_map <- workshops %>%
  filter(city != "Online")

# Geocode all locations using OpenStreetMap via tidygeocoder
workshops_geocoded <- workshops_for_map %>%
  geocode(Location, method = 'osm', lat = lat, long = lon)

```

```{r}
# Count total number of workshops (excluding NA in the date column)
total_workshops <- workshops |> filter(!is.na(date)) |> nrow()

# Display the total number of workshops with HTML styling
htmltools::HTML(
  paste0(
    "<div class='workshop-count' style='text-align: center; font-size: 1.2em; font-weight: bold; color: var(--default-body); margin-bottom: 20px; background-color: var(--neutral-n50); padding: 10px; border-radius: 5px;'>",
    "Total number of workshops: ", total_workshops, 
    "</div>"
  )
)

```

```{r}
# Define explicit coordinate limits for a consistent world view
lon_limits <- c(-180, 180)
lat_limits <- c(-60, 90)

# Prepare the world map data
world_map <- map_data("world") %>%
  filter(region != "Antarctica")

# Create the ggplot object for the map
gg <- ggplot() +
  geom_polygon(
    data = world_map,
    aes(x = long, y = lat, group = group),
    fill = "grey85",
    color = "white"
  ) +
  geom_jitter(
    data = workshops_geocoded,
    aes(
      x = lon,
      y = lat,
      text = paste0(
        "<b>Date:</b> ", date, "<br>",
        "<b>Hosted By:</b> ", location_name, "<br>",
        "<b>Location:</b> ", Location, "<br>",
        "<b>Workshop:</b> ", title, "<br>",
        "<b>Instructors:</b> ", instructors
      ),
      colour = title,
      fill = title,
      shape = title,
      size = title
    ),
    width = 0.5,
    height = 0.5
  ) +
  scale_colour_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2") +
  scale_shape_manual(values = 21:25) +  # Adjust number of shapes as needed
  scale_size_manual(values = rep(2, length(unique(workshops_geocoded$title)))) +
  coord_fixed(
    ratio = 1.3,
    xlim = lon_limits,
    ylim = lat_limits,
    expand = FALSE
  ) +
  theme_void() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.5, "lines"),
    legend.margin = margin(t = 0, b = 0, unit = 'pt')
  )

# Convert ggplot to interactive plotly plot
interactive_map <- ggplotly(gg, tooltip = "text") %>%
  layout(
    margin = list(l = 20, r = 20, t = 20, b = 50),
    hoverlabel = list(align = "left"),
    legend = list(
      orientation = "h",
      x = 0.5,
      xanchor = "center",
      y = -0.1
    ),
    xaxis = list(
      showline = FALSE,
      showticklabels = FALSE,
      zeroline = FALSE
    ),
    yaxis = list(
      showline = FALSE,
      showticklabels = FALSE,
      zeroline = FALSE
    )
  ) %>%
  config(
    responsive = TRUE,
    displaylogo = FALSE
  )

# Display the interactive map
interactive_map
```

The table below lists all workshops, including the date, host institution, city, country, and region. The region column has been added to make it easier to find workshops in the same geographical area. These regions are based on the World Bank's classification system. The table is fully searchable and filterable, allowing you to quickly locate workshops by various criteria, such as location or instructors.

```{r}
# Create a separate data frame for the table with HTML links
workshops_for_table <- workshops %>% 
  mutate(
    location_name_html = if_else(
      !is.na(location_url) & location_url != "",
      sprintf("<a href='%s' target='_blank'>%s</a>", location_url, location_name),
      location_name
    ),
    title_html = if_else(
      !is.na(workshop_url) & workshop_url != "",
      sprintf("<a href='%s' target='_blank'>%s</a>", workshop_url, title),
      title
    )
  )

# Create the reactable table
workshops_for_table %>% 
  select(date, location_name_html, title_html, city, country, region, instructors) %>% 
  reactable(
    filterable = TRUE,
    searchable = TRUE,
    compact = TRUE,
    columns = list(
      date = colDef(name = "Date"),
      location_name_html = colDef(name = "Hosted By", html = TRUE),
      title_html = colDef(name = "Workshop", html = TRUE),
      city = colDef(name = "City"),
      country = colDef(name = "Country"),
      region = colDef(name = "Region"),
      instructors = colDef(name = "Instructors")
    ),
    defaultPageSize = 100,
    striped = TRUE,
    bordered = TRUE,
    style = list(
      border = "1px solid #ccc",
      stripedColor = "#f0f0f0",
      fontSize = "0.875rem"
    )
  )
```

:::: footer-timestamp
::: {style="text-align: center; font-size: 0.9em; color: var(--neutral-n50); margin-top: 20px; padding-top: 10px;"}
```         
Last rendered on: `r format(Sys.Date(), "%Y-%m-%d")`
```
:::
::::
